---
title: "Building molecular networks and patient-patient similarity networks"
author: '[Marianyela Petrizzelli](mailto:marianyela.petrizzelli@curie.fr)'
date: "January 2021"
output:
  md_document:
    toc: yes
    toc_depth: 2
  html_document:
    depth: 3
    keep_md: yes
    number_sections: yes
    toc: yes
---
This package provides the codes to Build Molecular Networks and Patient-Patient Similarity networks (BMNPPS) described in details in deliverable 4.1. 
The package is currently being tested.

# Intalling BMNPPS
The BMNPPS package relies on R version 4.0.3, thus BioConductor must be updated to version 3.12. 

```{r, eval=TRUE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.12")
```
In addition, BMNPPS package relies on package `gprofiler2` that can be installed using `BiocManager`, and `rROMA` that can be installed using `devtools`:

```{r, eval=FALSE}
if(!"gprofiler2" %in% installed.packages()){
  BiocManager::install("gprofiler2")
}
if(!requireNamespace("devtools")){
  install.packages("devtools")
}
if(!requireNamespace("rRoma")){
  devtools::install_github("Albluca/rROMA")
}
```
Similarly, it is possible to install the BMNPPS package using devtools by typing:
```{r, eval=FALSE}
if(!requireNamespace("BMNPPS")){
  devtools::install_github("iPC-project-H2020/wp4-deliverables/BMNPPS")
}
```

# Using BMNPPS
Upload the BMNPPS package:
```{r setup}
library(BMNPPSN)
```
BMNPPS requires an expression matrix - with column names indicating samples and row names indicating gene names - and a network file produced by COSIFER - with columns e1, e2 and intensity, in which each row indicates the intensity of the edge between node e1 and node e2. 

# An example

