---
title: "Building molecular networks and patient-patient similarity networks"
author: '[Marianyela Petrizzelli](mailto:marianyela.petrizzelli@curie.fr)'
date: "January 2021"
output:
  md_document:
    toc: yes
    toc_depth: 2
  html_document:
    depth: 3
    keep_md: yes
    number_sections: yes
    toc: yes
---
This package provides the codes to Build Molecular Networks and Patient-Patient Similarity Networks (BMNPPSN) described in details in deliverable 4.1. 
The package is currently being tested.

# Intalling BMNPPSN
The BMNPPSN package relies on R version 4.0.3, thus BioConductor must be updated to version 3.12. 

```{r, eval=TRUE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.12")
```
In addition, BMNPPSN package relies on package `gprofiler2` that can be installed using `BiocManager`, and `rROMA` that can be installed using `devtools`:

```{r, eval=FALSE}
if(!"gprofiler2" %in% installed.packages()){
  BiocManager::install("gprofiler2")
}
if(!requireNamespace("devtools")){
  install.packages("devtools")
}
if(!requireNamespace("rRoma")){
  devtools::install_github("Albluca/rROMA")
}
```
Similarly, it is possible to install the BMNPPS package using devtools by typing:
```{r, eval=FALSE}
if(!requireNamespace("BMNPPSN")){
  devtools::install_github("iPC-project-H2020/wp4-deliverables/BMNPPSN")
}
```

# Using BMNPPSN
Upload the BMNPPSN package:
```{r setup}
library(BMNPPSN)
```
BMNPPS requires an expression matrix - with column names indicating samples and row names indicating gene names - and a network file produced by COSIFER - with columns e1, e2 and intensity, in which each row indicates the intensity of the edge between node e1 and node e2. 

# Create a subnetwork
It is possible to create a subnetwork simply by typing 
`create_subnetwork(directory, outDirectory, n.edges=10^5)`
where `directory` is the directory of the original network in .csv.gz format, and `outDirectory` the user defined output directory. 

# Tunning the inflation parameter for clustering and clustering
Tunning and clustering is as simple as typing `mcl_clustering(resDirectory, outDirectory)` where now `resDirectory` is the directory where subnetworks have been saved and `outDirectory` the user defined output directory. MCL clustering is performed with inflation parameters from 1 to 3.5 with a step of 0.1. The inflation for which the max number of clusters with more than 10 genes is attaint, is retained and stored. 

# Enrichment analysis of the communities identified through MCL
Now that we have identify communities in the network, `run_gprofiler(directory)` performs gene enrichment analysis of the communities (saved in `directory`) and stores the results in `directory`. 

# Construction of eigenGene matrix using Roma
To construct the eigenGene matrix we need now an input `directory` where the expression matrix is stored and the `outDirectory` containing the results from the previous analysis. Thus, `runRoma(directory, outDirectory)`computes the eigenGene matrix and stores it in the `outDirectory`. 

# Knn patient similarity networks
Once we have the eigenGene matrix, construction of knn graph can be performed just by typing:`knn.mat <- nng(x = eigenGene.matrix, k=sel.k, method = "euclidean", mutual = TRUE)`, where `sel.k` is the user selected k in the knn graph. 

# Knn patient for WP4.1 multi-omics datasets
The construction of knn graph for the multiomics datasets considered in WP4.1 is made available in folder `R\pca_knn_graph.R` and the eigenGene matrices in `data` 
      
